---
title: "Figuring out how to test"
output: html_document
date: "2024-04-17"
---

# Background
Testing apps is hard and I don't know how to do it or how to even begin....


# The Beginning...
The first thing I want to test is the first page of my app, the file upload. To test this I think I need to figure out how to set up a testServer. I'm going to make a small dummy app here to test this.

```{r}
ui <- shiny::fluidPage(
  shiny::fileInput("file_input", "File Input")
)

server <- function(input, output, session){
  file <- reactive({
    req(input$file_input)
    readRDS(input$file_input$datapath)
  })
}

testServer(
  app = server,
  args = list(),
  exp = {
    ns <- session$ns
    expect_true(is.null(input$file_input$datapath))
    session$setInputs(file_input = list(datapath = "data/example_rds.rds"))
    expect_true(!is.null(input$file_input))
    expect_true(is.data.frame(file()))
  }
)
```

Note how I have input file_input as a list with the entry named datapath. This is to imitate how the `shiny::fileInput` function receives input files in an interactive context.

The big thing I don't know how to test is how the ui changes in response to reactivity in the server. For example, the default ui value has no file selected by the fileInput widget:

```{r}
ui_char <- as.character(ui)
expect_true(stringr::str_detect(ui_char, "No file selected"))
```

But how do I check that this changes if a file is selected? I suspect I am being pedantic here - is it shiny's responsability to make sure that `fileInput` actually uploads a fileInput... this seems sensible. Maybe I should concern myself with the other 100 things I don't know how to do.

## Trying some things from the shiny book

Here is a ui and server function:
```{r}
summaryUI <- function(id) {
  tagList(
    outputText(ns(id, "min")),
    outputText(ns(id, "mean")),
    outputText(ns(id, "max")),
  )
}
summaryServer <- function(id, var) {
  stopifnot(is.reactive(var))
  
  moduleServer(id, function(input, output, session) {
    range_val <- reactive(range(var(), na.rm = TRUE))
    output$min <- renderText(range_val()[[1]])
    output$max <- renderText(range_val()[[2]])
    output$mean <- renderText(mean(var()))
  })
}

```

Here are some tests:
```{r}
x <- reactiveVal(1:10)
testServer(summaryServer, args = list(var = x), {
  print(range_val())
  print(output$min)
})
#> [1]  1 10
#> [1] "1"
#> 

test_that("output updates when reactive input changes", {
  x <- reactiveVal()
  testServer(summaryServer, args = list(var = x), {
    x(1:10)
    browser()
    session$flushReact()
    expect_equal(range_val(), c(1, 10))
    expect_equal(output$mean, "5.5")
    
    x(10:20)
    session$flushReact()
    expect_equal(range_val(), c(10, 20))
    expect_equal(output$min, "10")
  }) 
})
#> Test passed 🌈
```

Ok so here we use `session$flushReact()` because the outputs don't update the server outputs even though they do update the reactive `range_val` (try it with `session$flushReact` commented out and see what happens). This is allegedly because x is defined externally to the server. I wonder can I define it internally or will this not work as summaryServer requires it as an input?

```{r}
test_that("output updates when reactive input changes", {
  
  testServer(summaryServer, args = list(), {
    x <- reactiveVal()
    x(1:10)
    browser()
    session$flushReact()
    expect_equal(range_val(), c(1, 10))
    expect_equal(output$mean, "5.5")
    
    x(10:20)
    session$flushReact()
    expect_equal(range_val(), c(10, 20))
    expect_equal(output$min, "10")
  }) 
})
```

No it fails because of the `stopifnot` functionality (duh!)!!!

## Some more detail
Okay so now I've kind of worked out how to test my server functions (this is the overstatement of the century):

1. when I add inputs, are the appropriate reactions triggered?
2. Don't test things that shiny should be testing
3. I still don't know how to test ui changes based on reactive server variables

But `clusteringServer` is still giving me grief. `clusteringServer` outputs a plotly graph which registers some event_data so that I can do other cool(?) things, but that `shiny::event_data` call gives a very annoying warning message if it is called before there is any data available in the graph. I have avoided this warning message when running the app itself by using things like `shiny::req()` before evaluating calls to `plotly::event_data` but for some reason, this warning still appears when I test the server and seems to mean that tests don't pass (this is just a warning rather than an error and I'm not using `expect_silent` so this too is a mystery).

Ok let's ignore the above for a minute because I've lost my train of thought and look at testing uploading the reduced embeddings. 

As an aside, I want to get to a point where I can just make a dummy df inside these testServers rather than uploading df and reduced_embeddings each time but for now I just want to get these test working and then I can start making them more efficient.

```{r}
testServer(
  app = clusteringServer,
  args = list(),
  exp = {
    
    # inputs
     session$setInputs(
       data_upload = list(
         name = "data/example_rds.rds",
         datapath = "data/example_rds.rds"
         ),
       load_or_reduce_embeddings = "Load in reduced embeddings",
       reduced_embeddings_upload = 
         list(
           name = "inst/testdata/reduced_embeddings.csv",
           datapath = "inst/testdata/reduced_embeddings.csv"
           )
       )
    
    expect_true(inherits(reduced_embeddings(), "data.frame"))
  }
)
```

Ok that works, great! Does it work if I make a dummy df?

```{r}
testServer(
  app = clusteringServer,
  args = list(),
  exp = {
    
    # inputs
     session$setInputs(
       load_or_reduce_embeddings = "Load in reduced embeddings",
       reduced_embeddings_upload = 
         list(
           name = "inst/testdata/reduced_embeddings.csv",
           datapath = "inst/testdata/reduced_embeddings.csv"
           )
       )
    
    expect_true(inherits(reduced_embeddings(), "data.frame"))
  }
)
```

Yes ok, because uploading embeddings doesn't require a df.

But what about the more complicated task of calculating embeddings within the app. This is awkward because I require more than the `clusteringServer`, I also need the `reduceServer` and I need to be able to hit the `do_reducing_option1` action button.

```{r}
testServer(
  app = clusteringServer,
  args = list(id = "test"),
  exp = {
    ns <- session$ns
    # inputs to the clusteringServer
     session$setInputs(
       data_upload = list(
         name = "data/example_rds.rds",
         datapath = "data/example_rds.rds"
         ),
       load_or_reduce_embeddings = "Calculate in app"
       )
    
    # inputs to the reduceServer
    session$setInputs(
      `reduction_ui-n_neighbours1` = 10,
      `reduction_ui-n_components1` = 5,
      `reduction_ui-min_dist1` = 0.0,
      `reduction_ui-reducing_metric1` = "cosine",
      `reduction_ui-do_reducing_option1` = 1
      
    )
    
    browser()
    # expect_true(inherits(reduced_embeddings(), "data.frame"))
  }
)
```

An example from the internet:

```{r}
outerModule <- function(id) {
  moduleServer(id, function(input, output, session) {
    r1 <- reactive({ input$x + 1})
    r2 <- innerModule("inner", r1)
    output$someVar <- renderText(r2())
  })
}

innerModule <- function(id, r) {
  moduleServer(id, function(input, output, session) {
    reactive(paste("a value:", r()))
  })
}

testServer(outerModule, args = list(id = "foo"), {
  session$setInputs(x = 1)
  expect_equal(output$someVar, "a value: 2")
  browser()
})
```

Since I can't make this work maybe I should start with testing the `reducingServer` in isolation.

```{r}
df <- reactive({
    data.frame(docs = "this is a doc",
               v1 = 0.1,
               v2 = 0.3) %>%
      dplyr::mutate(embeddings = matrix(runif(5), nrow = 1))
}) 


testServer(
  app = reducingServer,
  args = list(id = "test",
              df = df),
  exp = {
    library(BertopicR)
    # inputs to the reduceServer
    session$setInputs(
      n_neighbours1 = 10,
      n_components1 = 5,
      min_dist1 = 0.0,
      reducing_metric1 = "cosine",
      do_reducing_option1 = 1
    )
     Sys.sleep(10)
     print(reduced_embeddings1$get_result())
    # browser()
    # expect_true(inherits(reduced_embeddings(), "data.frame"))
  }
)
```

Still not working - am I missing something obvious? Let's testing background_reduce:

```{r}
# df_upload <- readr::read_rds("data/example_rds.rds")
df_test <- data.frame(docs = "this is a doc",
               v1 = 0.1,
               v2 = 0.3) %>%
      dplyr::mutate(embeddings = matrix(runif(5), nrow = 1))
testServer(
  app = backgroundReduce,
  args = list( 
    id = "test",
    n_neighbours = 10,
    n_components = 5,
    min_dist = 0.0,
    metric = "cosine",
    embeddings = df_test$embeddings,
    wait_for_event = TRUE),
  exp = {
    
    # wait_for_event means job is not triggered yet
    expect_false(trigger_job())
    
    # not triggered yet
    expect_error(inherits(bg_job(), "r_process"))
    
    trigger_job(TRUE) # mimics action button in reduceServer
    expect_true(inherits(bg_job(), "r_process")) # job created
    expect_true(bg_job()$is_alive()) # running
    Sys.sleep(7) # wait for job to run
    expect_false(bg_job()$is_alive()) # finished running
    expect_true(is.array(bg_job()$get_result())) # outputs matrix
    
  }
)

```

Ok so the `backgroundReduce` function does in fact work. The issue I'm having is testing parent server functions because triggering the action button that changes `trigger_job` to TRUE does not seem to be working as intended when triggered from a `testServer`.


For the purpose of figuring out how to test this, I am going to return the `bg_job()$is_alive()` variable from the `backgroundReduce` function for now, I will remove it once I've figured this out, so if you're reading this later, this chunk should not work.
```{r}
df <- reactive({
    data.frame(docs = "this is a doc",
               v1 = 0.1,
               v2 = 0.3) %>%
      dplyr::mutate(embeddings = matrix(runif(5), nrow = 1))
}) 


testServer(
  app = reducingServer,
  args = list(id = "test",
              df = df),
  exp = {
    # library(BertopicR)
    # inputs to the reduceServer
    session$setInputs(
      n_neighbours1 = 10,
      n_components1 = 5,
      min_dist1 = 0.0,
      reducing_metric1 = "cosine",
      do_reducing_option1 = NULL,
      `backgroundReduce-n_neighbours` = 10,
      `backgroundReduce-n_components` = 5,
      `backgroundReduce-min_dist` = 0.0,
      `backgroundReduce-metric` = "cosine",
      `backgroundReduce-embeddings` = df()$embeddings,
      `backgroundReduce-wait_for_event` = TRUE,
    )
     
    browser()
    
    testfunc <- function(){
      session$setInputs(do_reducing_option1 = 1)
      session$elapse(millis = 100)
    }
    # session$setInputs(do_reducing_option1 = 1)
    # session$elapse(millis = 10000)
    # print(reduced_embeddings1$get_result())
     # Sys.sleep(10)
     # browser()
     print(reduced_embeddings1$get_result())
    # browser()
    # expect_true(inherits(reduced_embeddings(), "data.frame"))
  }
)

```

It is not obvious from the above code but basically the above works if I run `session$setInputs(do_reducing_option1 = 1)` and `session$elapse(millis = 10000)` concurrently in the console (ie. if I highlight both and press enter or type both into the console and type enter) but it does not work if I run them one after the other. This indicates that there is some timing issue related to the `invalidateLater` function in `backgroundReduce`. The `invalidateLater` timing had previously been 250ms but which I upped to 2500ms to see if it solved the issue - it did not. 

**I made this work in mt test-Reduce file. If I don't update here check that!**

```{r}
df <- function(){
    dummy_df <- data.frame(docs = "this is a doc",
               # embeddings = embeddings,
               v1 = 0.1,
               v2 = 0.3)

    return(list(df= dummy_df))
}

testServer(
  app = clusteringServer,
  args = list(),
  exp = {
    # ns <- session$ns
    df <- reactiveVal()
    df(data.frame(docs = "this is a doc",

              v1 = 0.1,
              v2 = 0.3) %>%
       dplyr::mutate(embeddings = matrix(runif(5), nrow = 1)))
    browser()
    expect_true(is.data.frame(df()))
    #

    # session$setInputs(file_input = list(datapath = "data/example_rds.rds"))
    # expect_true(!is.null(input$file_input))
    # expect_true(is.data.frame(file()))
  }
)

test_that("Calculate in app option calculates reduced embeddings", {
  testServer(
    app = clusteringServer,
    args = list(),
    exp = {
      session$setInputs(data_upload =
                          list(
                            name = "data/example_rds.rds",
                            datapath = "data/example_rds.rds"),
                            load_or_reduce_embeddings = "Calculate in app")
      session$setInputs(`reduction_ui-df` = df,)
      browser()
      # expect_true(!is.null(reduced_embeddings()))
    }
  )
})
```

The following has been copied nad pasted from the interenet and I am just using it for testing:

```{r}

library(shiny)
library(testthat)

test_that("Nested modules", {
  child <- function(id) {
    moduleServer(id, function(input, output, session) {
      output$txt <- renderText("bar")
    })
  }

  parent <- function(id) {
    moduleServer(id, function(input, output, session) {
      output$txt <- renderText("foo")
      child("child-id")
    })
  }

  testServer(parent, args = list(id = "parent-id"), {
    browser()
    expect_equal(output$txt, "foo")
  })

})

test_that("Lack of ID", {
  module <- function(id) {
    moduleServer(id, function(input, output, session) {
      output$txt <- renderText(session$ns("x"))
    })
  }

  testServer(module, args = list(id = "foo"), {
    expect_equal(output$txt, "foo-x")
  })
})

test_that("testServer works with nested module servers", {
  outerModule <- function(id) {
    moduleServer(id, function(input, output, session) {
      r1 <- reactive({ input$x + 1})
      r2 <- innerModule("inner", r1)
      output$someVar <- renderText(r2())
    })
  }

  innerModule <- function(id, r) {
    moduleServer(id, function(input, output, session) {
      reactive(paste("a value:", r()))
    })
  }

  testServer(outerModule, args = list(id = "foo"), {
    browser()
    session$setInputs(x = 1)
    
    r2()
    expect_equal(output$someVar, "a value: 2")
  })
})

test_that("testServer calls do not nest in module functions", {
  module <- function(id) {
    moduleServer(id, function(input, output, session) {
      x <- 1
      testServer(function(id) {
        moduleServer(id, function(input, output, session) {
          y <- x + 1
        })
      })
    })
  }

  expect_error(testServer(module, {}))
})

test_that("testServer calls do not nest in test exprs", {
  module <- function(id) {
    x <- 1
    moduleServer(id, function(input, output, session) {
      inner <- function(id) {
        moduleServer(id, function(input, output, session) {
          y <- x + 1
        })
      }
    })
  }

  expect_error(testServer(module, { testServer(inner, {}) }))
})
```
